---
title: "Normal Approximation to Binomial Distributions"
output: html_document
---

```{r include=FALSE}
library(ggplot2)
library(gridExtra)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Binomial distribution with n= 100

The sampling distribution of the binomial variable following $Bin(n=100, \pi = 0.01)$ has its mean and variance close to their theoretical values, but the shape does not look like the normal distribution.

The overlayed curve is the normal distribution with the theoretical mean and variance in case that the normal approximation applies.

```{r message = FALSE, fig2, fig.height = 3, fig.width = 6, fig.align = "center"}
n = 100
p = 0.01
d.binom <- data.frame(index=1:1000, events=rbinom(1000, n, p))

plot1 <- ggplot(d.binom, aes(x=events)) + geom_histogram(binwidth = 1)  + ggtitle(paste("histogram n=",n, "\n")) + theme(plot.title = element_text(size=10))

plot2 <- ggplot(d.binom, aes(x=events)) + geom_histogram(aes(y=..density..)) + stat_function(fun = dnorm, args = list(mean = 1, sd = sqrt(100*0.01*(1-0.01)))) + ggtitle(paste("density n=",n, ", np=", n*p ,",\nmean=", round(mean(d.binom$events),1), ", variance=",round(var(d.binom$events),1))) + theme(plot.title = element_text(size=10))

grid.arrange(plot1, plot2, ncol=2)

```


The mean of the distribution is the same as its variance, which is one of the properties of the Poisson distribution. Since the number of iterations is very large, the sampling distribution of the binomial distribution becomes the poisson distribution.

```{r fig3, fig.height = 3, fig.width = 3, fig.align = "center"}
d.pois <- data.frame(events = rpois(1000,lambda = sum(d.binom$events)/nrow(d.binom)))

ggplot(d.pois, aes(x=events)) + geom_histogram(binwidth = 1) + ggtitle(paste("the poisson dist, \nmean=", round(mean(d.pois$events),1), ", variance=",round(var(d.pois$events),1)))+ theme(plot.title = element_text(size=10))


```

## Binomial distributions with difefrent sample sizes

### n = 500
```{r}
n = 500
p = 0.01
d.binom <- data.frame(index=1:1000, events=rbinom(1000, n, p))

plot500 <- ggplot(d.binom, aes(x=events)) + geom_histogram(aes(y=..density..)
) + stat_function(fun = dnorm, args = list(mean = 5, sd = sqrt(500*0.01*(1-0.01)))) + ggtitle(paste("density n=",n, ", np=", n*p,", \nmean=", round(mean(d.binom$events),1), ", variance=",round(var(d.binom$events),1))) + theme(plot.title = element_text(size=10))


```

### n = 1000
```{r}
n = 1000
p = 0.01
d.binom <- data.frame(index=1:1000, events=rbinom(1000, n, p))

plot1000 <- ggplot(d.binom, aes(x=events)) + geom_histogram(aes(y=..density..)
) + stat_function(fun = dnorm, args = list(mean = 10, sd = sqrt(1000*0.01*(1-0.01))))+ ggtitle(paste("density n=",n, ", np=", n*p,", \nmean=", round(mean(d.binom$events),1), ", variance=",round(var(d.binom$events),1))) + theme(plot.title = element_text(size=10))

```


### n = 2000
```{r}
n= 2000
p = 0.01
d.binom <- data.frame(index=1:100, events=rbinom(1000, 10000, 0.01))

plot2000 <- ggplot(d.binom, aes(x=events)) + geom_histogram(aes(y=..density..)
) + stat_function(fun = dnorm, args = list(mean = 100, sd = sqrt(10000*0.01*(1-0.01)))) + ggtitle(paste("density n=",n, ", np=", n*p,", \nmean=", round(mean(d.binom$events),1), ", variance=",round(var(d.binom$events),1))) + theme(plot.title = element_text(size=10))

```


### 4 Plots
Here is the set of 4 plots generated with different binomial distributions with different sample sizes. It is said that the condition in which we can use the normal approximation is $np \ge 10$, and it is confirmed with the plot with $np = 10$ below, which looks normal.

```{r message=FALSE, fig4, fig.height = 6, fig.width = 6, fig.align = "center"}
grid.arrange(plot2, plot500, plot1000, plot2000, ncol=2)
```

